### 2021-08-02 supplement.tex
all figures

\section{Supplementary Material}
\beginsupplement

\begin{figure}[H]
  \centering
  \fbox{
   \includegraphics[width=0.36\paperwidth]{fig/hcp_iq_nonnorm_hist.pdf}
   \includegraphics[width=0.36\paperwidth]{fig/hcp_iq_quanttrf_hist.pdf}
   }
  \caption{Histogram of fluid intelligence score in the HPC dataset, before (left) and after (right) quantile transformation.}
  \label{fig:hcp-hist}
\end{figure}

\begin{figure}[H]
  \centering
  \fbox{
   \includegraphics[width=0.36\paperwidth]{fig/abide_motion_hist.pdf}
   \includegraphics[width=0.36\paperwidth]{fig/abide_motion_quanttrf_hist.pdf}
   }
  \caption{Histogram of mean framewise displacement in the ABIDE dataset, before (left) and after (right) quantile transformation.}
  \label{fig:abide-hist}
\end{figure}