### 2021-10-20 fig/raw/emp_abide_motion_raw_partial.tex

\documentclass{article}
\usepackage[x11names, svgnames, rgb]{xcolor}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usetikzlibrary{snakes,arrows,shapes}
\usepackage{amsmath}
%
%

\usepackage[active,tightpage]{preview}
\PreviewEnvironment{tikzpicture}
\setlength\PreviewBorder{0pt}%

%

\begin{document}
\pagestyle{empty}
%
%
%

\enlargethispage{100cm}
% Start of code
\begin{tikzpicture}[>=latex',line join=bevel,]
%%
\node (c) at (27.0bp,18.0bp) [draw,ellipse] {c};
  \node (y) at (197.5bp,53.0bp) [draw,ellipse] {y};
  \node (yhat) at (368.0bp,18.0bp) [draw,ellipse] {<<y\&\#770;>>};
  \draw [] (c) ..controls (85.084bp,29.857bp) and (139.57bp,41.174bp)  .. (y);
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (92.5bp,42.5bp) node {0.028};
  \draw [dashed] (c) ..controls (74.615bp,14.069bp) and (104.67bp,11.894bp)  .. (131.0bp,11.0bp) .. controls (190.08bp,8.9948bp) and (204.92bp,8.9948bp)  .. (264.0bp,11.0bp) .. controls (290.33bp,11.894bp) and (320.39bp,14.069bp)  .. (yhat);
  \draw (197.5bp,18.5bp) node {0.111 (p<0.0001*)};
  \draw [] (y) ..controls (255.58bp,41.143bp) and (310.07bp,29.826bp)  .. (yhat);
  \draw (302.5bp,42.5bp) node {0.126};
%
\end{tikzpicture}
% End of code

%
\end{document}
%