### 2021-10-20 fig/raw/emp_abide_motion_reg_full.tex

\documentclass{article}
\usepackage[x11names, svgnames, rgb]{xcolor}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usetikzlibrary{snakes,arrows,shapes}
\usepackage{amsmath}
%
%

\usepackage[active,tightpage]{preview}
\PreviewEnvironment{tikzpicture}
\setlength\PreviewBorder{0pt}%

%

\begin{document}
\pagestyle{empty}
%
%
%

\enlargethispage{100cm}
% Start of code
\begin{tikzpicture}[>=latex',line join=bevel,]
%%
\node (c) at (27.0bp,18.0bp) [draw,ellipse] {c};
  \node (y) at (158.0bp,53.0bp) [draw,ellipse] {y};
  \node (yhat) at (339.0bp,18.0bp) [draw,ellipse] {<<y\&\#770;>>};
  \draw [] (c) ..controls (75.327bp,30.841bp) and (109.82bp,40.199bp)  .. (y);
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (92.5bp,47.5bp) node {0.028};
  \draw [] (c) ..controls (74.615bp,14.069bp) and (104.67bp,11.894bp)  .. (131.0bp,11.0bp) .. controls (196.05bp,8.7922bp) and (272.59bp,13.206bp)  .. (yhat);
  \draw (158.0bp,18.5bp) node {0.359};
  \draw [dashed] (y) ..controls (218.38bp,41.389bp) and (278.81bp,29.573bp)  .. (yhat);
  \draw (248.5bp,50.5bp) node {0.02 (p=0.1)};
%
\end{tikzpicture}
% End of code

%
\end{document}
%