### 2021-10-21 simulated/normality_and_linearity_violation.tex
example notebook

\hypertarget{an-example-of-model-non-linearity-in-the-presence-of-a-linear-predictor.}{%
\section{An example of model non-linearity in the presence of a linear
predictor.}\label{an-example-of-model-non-linearity-in-the-presence-of-a-linear-predictor.}}

Notebook available on-line:
https://github.com/spisakt/mlconfound\_manuscript/tree/main/simulated/normality\_and\_linearity\_violation.ipynb

\begin{Shaded}
\begin{Highlighting}[]
\ImportTok{import}\NormalTok{ numpy }\ImportTok{as}\NormalTok{ np}
\ImportTok{import}\NormalTok{ seaborn }\ImportTok{as}\NormalTok{ sns}
\ImportTok{import}\NormalTok{ matplotlib.pyplot }\ImportTok{as}\NormalTok{ plt}
\ImportTok{from}\NormalTok{ sklearn.linear_model }\ImportTok{import}\NormalTok{ Ridge}
\NormalTok{sns.}\BuiltInTok{set}\NormalTok{(rc}\OperatorTok{=}\NormalTok{\{}\StringTok{"figure.figsize"}\NormalTok{:(}\DecValTok{6}\NormalTok{, }\DecValTok{2}\NormalTok{)\})}
\NormalTok{sns.set_style(}\StringTok{"white"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{five-features-one-linear-four-sigmoid}{%
\subsubsection{Five features: one linear, four
sigmoid}\label{five-features-one-linear-four-sigmoid}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{n}\OperatorTok{=}\DecValTok{300}
\NormalTok{p}\OperatorTok{=}\DecValTok{5}
\NormalTok{rng }\OperatorTok{=}\NormalTok{ np.random.default_rng(}\DecValTok{42}\NormalTok{)}
\NormalTok{y }\OperatorTok{=}\NormalTok{ np.arange(n)}
\NormalTok{y }\OperatorTok{=}\NormalTok{ (y }\OperatorTok{-}\NormalTok{ y.mean())}\OperatorTok{/}\NormalTok{y.std()}

\NormalTok{X_true }\OperatorTok{=}\NormalTok{ np.repeat(y, p).reshape(n,p)}
\ControlFlowTok{for}\NormalTok{ i }\KeywordTok{in} \BuiltInTok{range}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{4}\NormalTok{):}
\NormalTok{    X_true[:,i] }\OperatorTok{=}\NormalTok{ np.tanh(X_true[:,i]}\OperatorTok{*}\DecValTok{2}\NormalTok{)}

\NormalTok{X}\OperatorTok{=}\NormalTok{X_true }\OperatorTok{+}\NormalTok{ rng.normal(}\DecValTok{0}\NormalTok{,}\FloatTok{0.1}\NormalTok{, (n,p))}

\ControlFlowTok{for}\NormalTok{ i }\KeywordTok{in} \BuiltInTok{range}\NormalTok{(}\DecValTok{4}\NormalTok{):}
\NormalTok{    sns.lineplot(x}\OperatorTok{=}\NormalTok{y, y}\OperatorTok{=}\NormalTok{X_true[:,i])}
\NormalTok{    sns.scatterplot(x}\OperatorTok{=}\NormalTok{y, y}\OperatorTok{=}\NormalTok{X[:,i])}
\NormalTok{    plt.show()}
\end{Highlighting}
\end{Shaded}

\begin{figure}
\centering
\includegraphics{normality_and_linearity_violation_files/normality_and_linearity_violation_3_0.png}
\caption{png}
\end{figure}

\begin{figure}
\centering
\includegraphics{normality_and_linearity_violation_files/normality_and_linearity_violation_3_1.png}
\caption{png}
\end{figure}

\begin{figure}
\centering
\includegraphics{normality_and_linearity_violation_files/normality_and_linearity_violation_3_2.png}
\caption{png}
\end{figure}

\begin{figure}
\centering
\includegraphics{normality_and_linearity_violation_files/normality_and_linearity_violation_3_3.png}
\caption{png}
\end{figure}

\hypertarget{model-predictions-without-regularization}{%
\subsubsection{\texorpdfstring{Model predictions \textbf{without}
regularization}{Model predictions without regularization}}\label{model-predictions-without-regularization}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{model }\OperatorTok{=}\NormalTok{ Ridge(alpha}\OperatorTok{=}\DecValTok{0}\NormalTok{)}
\NormalTok{model.fit(y}\OperatorTok{=}\NormalTok{y, X}\OperatorTok{=}\NormalTok{X)}
\NormalTok{yhat }\OperatorTok{=}\NormalTok{ model.predict(X)}
\NormalTok{sns.scatterplot(x}\OperatorTok{=}\NormalTok{y, y}\OperatorTok{=}\NormalTok{yhat, alpha}\OperatorTok{=}\FloatTok{0.3}\NormalTok{)}
\CommentTok{#sns.lineplot(x=y, y=y, linestyle="--")}
\NormalTok{model.coef_}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
array([ 0.47902429,  0.00946499,  0.07378478, -0.02828672,  0.47070535])
\end{verbatim}

\begin{figure}
\centering
\includegraphics{normality_and_linearity_violation_files/normality_and_linearity_violation_5_1.png}
\caption{png}
\end{figure}

\hypertarget{model-predictions-with-regularization}{%
\subsubsection{\texorpdfstring{Model predictions \textbf{with}
regularization}{Model predictions with regularization}}\label{model-predictions-with-regularization}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{model }\OperatorTok{=}\NormalTok{ Ridge(alpha}\OperatorTok{=}\DecValTok{2000}\NormalTok{)}
\NormalTok{model.fit(y}\OperatorTok{=}\NormalTok{y, X}\OperatorTok{=}\NormalTok{X)}
\NormalTok{yhat }\OperatorTok{=}\NormalTok{ model.predict(X)}
\NormalTok{sns.scatterplot(x}\OperatorTok{=}\NormalTok{y, y}\OperatorTok{=}\NormalTok{yhat, alpha}\OperatorTok{=}\FloatTok{0.3}\NormalTok{)}
\CommentTok{#sns.lineplot(x=y, y=y, linestyle="--")}
\NormalTok{model.coef_}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
array([0.09472368, 0.07472972, 0.07338482, 0.07440682, 0.09463765])
\end{verbatim}

\begin{figure}
\centering
\includegraphics{normality_and_linearity_violation_files/normality_and_linearity_violation_7_1.png}
\caption{png}
\end{figure}

\begin{Shaded}
\begin{Highlighting}[]

\end{Highlighting}
\end{Shaded}